\textbf{Discovery of a disrupting open cluster in the Galactic halo:\\ a recent star formation event in the leading arm of the Magellanic stream}

### 2018-11-03 ms.tex
authors
reorg

We report the discovery of a young $\tau \sim \clage$, metal-poor $[\textrm{Fe}/\textrm{H}] \sim \clfeh$, low-mass $M \sim 1200~\msun$ stellar association in the Milky Way halo, Galactocentric $(R, z) \sim (23, 15)~\textrm{kpc}$.
The association --- \clustername\ --- is spatially near the leading arm of the gas stream emanating from the Magellanic cloud system, but is located $\approx 60^\circ$ from the Large Magellanic Cloud (LMC) center, on the other side of the Milky Way disk relative to the LMC.
By assuming that the cluster is co-located with \hi gas in the stream, we directly measure the distance to the Magellanic stream, $D \approx \cldist$.
At this location relative to the LMC, the measured distance is inconsistent with predictions from models of the LMC/SMC interaction and infall into the Milky Way that do not account for ram pressure and gas interaction with MW disk.
The estimated age of \clustername\ is consistent with the time of last passage through the Galactic midplane, and we therefore speculate that this star-formation event was triggered by the last disk midplane passage of the Magellanic stream gas, which occurred at a Galactocentric radius $R \approx 30~\kpc$.
However, most details of this idea remain a mystery: the Magellanic stream is low column density, the Milky Way disk at this large radius has low surface density, and the relative velocity of the stream gas and MW gas is large.
However it formed, the discovery of a young stellar cluster in the Milky Way halo presents an interesting opportunity for study.
\end{abstract}

@@

The halo of the Milky Way is characterized by its old ($\gtrsim 10~\textrm{Gyr}$), metal-poor ($\feh \approx -1.5$) stellar population.
This is understood as a signature of the dominant (in stellar mass) progenitor systems that were accreted and disrupted early on in the formation of the Galaxy: massive dwarf galaxies \citep[e.g.,][]{Deason:2015, Fiorentino:2015}.
These systems likely came in with significant gas reservoirs, but were quenched and stripped through collisional processes that heated and dispersed the gas \citep[e.g.,][]{Mayer:2006}, thus preventing immediate star formation in the deposited gas.
The Milky Way, however, continues to accrete satellite galaxies, as is evidenced by the prominent stellar stream from the Sagittarius dwarf galaxy \citep{Ibata:1994, Majewski:2003}, the presence of the Large and Small Magellanic Clouds (LMC, SMC), and about 50 dwarf satellites within the virial radius of the Galaxy.
While Sagittarius was likely stripped of its neutral gas long ago through repeated passage about the Milky Way \citep{Burton:1999, Tepper-Garcia:2018}, the \lmcsmc\ system is associated with $\approx 8\times 10^8~\msun$ of \hi gas \citep{Bruns:2005}, which extends into leading and trailing gas streams that wrap nearly $\approx 200^\circ$ around the sky \citep{Mathewson:1974, Putman:1998, Bruns:2005, Nidever:2010}.

The Magellanic stream (MS) is a large stream of (predominantly) hydrogen gas emanating from the \lmcsmc\ system that contains a significant fraction of the total gas mass associated with the \lmcsmc\ \citep{Bruns:2005}.
The stream was discovered in early surveys of 21 cm emission and has since been studied in great detail by large-area radio sky surveys, and in H$\alpha$ emission \citep{Weiner:1996, Barger:2017}.
With increased resolution, surveys have found small-scale structure and gas fragmentation \citep[e.g.,][]{Nidever:2008, For:2014} and a large-scale bifurcation, with %semi-distinct ``strands'' that lead back to the LMC and broader system, respectively \citep[e.g.,][]{Morras:1983, Putman:2003}.
kinematically \citep{Nidever:2008} and chemically \citep{Fox:2013} distinct ``strands'' that lead back to the LMC and the SMC, respectively.
As surveys push deeper to lower column densities, extensions of the stream have been found both in the trailing region \citep{Nidever:2010}, but also in the leading arm, which has been found to connect to regions of low-column-density gas ($N\sim 10^{18}$--$10^{19}~\textrm{cm}^{-2}$) on the other side of the Galactic disk \citep{Putman:1998, Nidever:2008}.
The leading arm (LA) has been decomposed into three or four distinct ``features'' named LA1--4 \citep{Bruns:2005, Nidever:2008, Venzmer:2012} with total HI mass $\sim 4 \times10^7~\msun$ \citep{Venzmer:2012}.

@@

We infer the mean proper motion of \clustername\ by constructing a probabilistic model of the cluster and background population using astrometric data from \gaia.
We start by selecting all stars with $\bprp_0 < 0.35$ to remove low-mass and old main sequence star contamination in the region.
\figurename~\ref{fig:pm-members}, left (grey points), shows the sky positions of stars that pass this blue cut in the region around the young cluster.
The larger solid-line circle indicates the region we define as the cluster area, and the two smaller dashed-line circles indicate control fields that are combined and used for modeling the background distribution of proper motions.
The control fields are designed to, together, have the same total area as the cluster field.
The right panel of \figurename~\ref{fig:pm-members} shows the difference of the 2D proper motion distributions in the cluster and summed control fields.
The distinct over-density of stars near $(\mu_\alpha, \mu_\delta) \approx (-0.5, 0.5)~\masyr$\footnote{Throughout this article, we use $\mu_\alpha$ to refer to the proper motion value provided by \gaia, which includes the $\cos\delta$ term.} is the identified comoving association of blue stars.

>>>>

We report the discovery of a young $\tau \sim \clage$, low-mass $M \sim 1200~\msun$, metal-poor $[\textrm{Fe}/\textrm{H}] \sim \clfeh$ stellar association in the Milky Way halo, \clustername.
At a Galactocentric position $(R, z) \sim (23, 15)~\textrm{kpc}$, the association is (on the sky) near the leading arm of the gas stream emanating from the Magellanic cloud system, but is located $\approx 60^\circ$ from the Large Magellanic Cloud (LMC) center, on the other side of the Milky Way disk relative to the LMC.
By assuming that the cluster is co-located with \hi gas in the stream, we directly measure the distance to the Magellanic stream, $D \approx \cldist$.
At this location relative to the LMC, the measured distance is inconsistent with predictions from models of the LMC/SMC interaction and infall into the Milky Way that do not account for ram pressure and gas interaction with MW disk.
The estimated age of \clustername\ is consistent with the time of last passage through the Galactic midplane, and we therefore speculate that this star-formation event was triggered by the last disk midplane passage of the Magellanic stream gas, which occurred at a Galactocentric radius $R \approx 30~\kpc$.
However, most details of this idea remain a mystery: the Magellanic stream is low column density, the Milky Way disk at this large radius has low surface density, and the relative velocity of the stream gas and MW gas is large.
However it formed, the discovery of a young stellar cluster in the Milky Way halo presents an interesting opportunity for study.
\end{abstract}

@@

% The halo of the Milky Way is characterized by its old ($\gtrsim 10~\textrm{Gyr}$), metal-poor ($\feh \approx -1.5$) stellar population.
% This is understood as a signature of the dominant (in stellar mass) progenitor systems that were accreted and disrupted early on in the formation of the Galaxy: massive dwarf galaxies \citep[e.g.,][]{Deason:2015, Fiorentino:2015}.
% These systems likely came in with significant gas reservoirs, but were quenched and stripped through collisional processes that heated and dispersed the gas \citep[e.g.,][]{Mayer:2006}, thus preventing immediate star formation in the deposited gas.
% The Milky Way, however, continues to accrete satellite galaxies, as is evidenced by the prominent stellar stream from the Sagittarius dwarf galaxy \citep{Ibata:1994, Majewski:2003}, the presence of the Large and Small Magellanic Clouds (LMC, SMC), and about 50 dwarf satellites within the virial radius of the Galaxy.
% While Sagittarius was likely stripped of its neutral gas long ago through repeated passage about the Milky Way \citep{Burton:1999, Tepper-Garcia:2018}, the \lmcsmc\ system is associated with $\approx 8\times 10^8~\msun$ of \hi gas \citep{Bruns:2005}, which extends into leading and trailing gas streams that wrap nearly $\approx 200^\circ$ around the sky \citep{Mathewson:1974, Putman:1998, Bruns:2005, Nidever:2010}.

The Magellanic stream (MS) is a large stream of (predominantly) hydrogen gas emanating from the \lmcsmc\ system that contains a significant fraction of the total gas mass associated with the \lmcsmc\ \citep{Bruns:2005} and wraps nearly $\approx 200^\circ$ around the sky \citep{Mathewson:1974, Putman:1998, Bruns:2005, Nidever:2010}.
The stream was discovered in early surveys of 21 cm emission and has since been studied in great detail by large-area radio sky surveys, and in H$\alpha$ emission \citep{Weiner:1996, Barger:2017}.
With increased resolution, surveys have found small-scale structure and gas fragmentation \citep[e.g.,][]{Nidever:2008, For:2014} and a large-scale bifurcation, with %semi-distinct ``strands'' that lead back to the LMC and broader system, respectively \citep[e.g.,][]{Morras:1983, Putman:2003}.
kinematically \citep{Nidever:2008} and chemically \citep{Fox:2013} distinct ``strands'' that lead back to the LMC and the SMC, respectively.
As surveys push deeper to lower column densities, extensions of the stream have been found both in the trailing region \citep{Nidever:2010}, but also in the leading arm, which has been found to connect to regions of low-column-density gas ($N\sim 10^{18}$--$10^{19}~\textrm{cm}^{-2}$) on the other side of the Galactic disk \citep{Putman:1998, Nidever:2008}.
The leading arm (LA) has been decomposed into three or four distinct ``features'' named LA1--4 \citep{Bruns:2005, Nidever:2008, Venzmer:2012} with total HI mass $\sim 4 \times10^7~\msun$ \citep{Venzmer:2012}.

@@

We infer the mean proper motion of \clustername\ by constructing a probabilistic model of the cluster and background populations using astrometric data from \gaia.
We start by selecting all stars with $\bprp_0 < 0.35$ to remove low-mass and old main sequence star contamination in the region.
\figurename~\ref{fig:pm-members}, left (grey points), shows the sky positions of stars that pass this blue cut in the region around the young cluster.
The larger solid-line circle indicates the region we define as the cluster area, and the two smaller dashed-line circles indicate control fields that are combined and used for modeling the background distribution of proper motions.
The control fields are designed to, together, have the same total area as the cluster field.
The right panel of \figurename~\ref{fig:pm-members} shows the difference of the 2D proper motion distributions in the cluster and summed control fields.
The distinct over-density of stars near $(\mu_\alpha, \mu_\delta) \approx (-0.5, 0.5)~\masyr$\footnote{Throughout this article, we use $\mu_\alpha$ to refer to the proper motion value provided by \gaia, which includes the $\cos\delta$ term.} is the identified comoving association of blue stars.
